---
output: 
  pdf_document:
    template: template.tex
---

```{r echo = FALSE, message = FALSE}
library(tidyverse)
library(bayesrules)
```

\section{Choice of Prior}
\vspace{10cm}

\section{Conjugate Prior}

Let $f(\theta)$ be a prior distribution for parmaeter $\theta$ from a particular distribution, and $L(\theta|x)$ be the likelihood function.
If the resulting posterior distribution $f(\theta|x)$ is of the same family of distributions as the prior distribution, then the prior distribution is a conjugate prior for this likelihood.

__Examples__

\newpage

\section{Non-conjugate prior for Binomial Data}

$$f(\pi)=e-e^\pi\; \text{ for } \pi \in [0,1] $$

Is this a valid pdf?


\vspace{5cm}


After observing 10 successes in 50 trials, calculate the posterior.

\newpage
\section{The Gamma-Poisson Conjugate Family}

\vspace{7cm}

\subsection{Gamma Prior}

Let $\lambda$ be a random variable which can take any value between 0 and $\infty$, ie. $\lambda \in [0,\infty)$.  Then the variability in $\lambda$ might be well modeled by a Gamma model with __shape parameter__ $\alpha > 0$ and __rate parameter__ 
$\beta > 0$: 

$\lambda \sim \text{Gamma}(\alpha, \beta)$

The Gamma distribution is specified by continuous pdf
\large$f(\lambda) = \frac{\beta^\alpha}{\Gamma(\alpha)} \lambda^{\alpha-1} e^{-\beta\lambda} \;\; \text{ for } \lambda \in [0,\infty)$


_Exercise_

What is the y-coordinate of the blue point in the plot below? In other words what is $f(\lambda)$ if $\lambda =1$?

```{r eval = FALSE}
plot_gamma(shape = 2, rate = 5)
```

```{r echo = FALSE, fig.align='center', fig.width=4, fig.height=3}
plot_gamma(2,5) +
    geom_point(aes(x = 1, y = dgamma(1, 
                                     shape =2, 
                                     rate =5)), colour="blue") +
  geom_text(aes(label= "(1, ?)", x = 1.25, y = 0.25), 
            color = "blue")
```


\newpage
__Descriptives of Gamma__

$E(\lambda) = \frac{\alpha}{\beta}$

$\text{Mode}(\lambda) = \frac{\alpha - 1}{\beta} \text{ where } \alpha\ge1$ 

$\text{Var}(\lambda) = \frac{\alpha}{\beta^2}$

__Gamma Tuning__

```{r echo=FALSE, gamma-tuning, fig.width = 4, fig.height = 3, message = FALSE, fig.align='center'}
# Set up gamma data
alpha <- c(2,2,2,   3,3,3,   5,5,5)
beta  <- c(2,4,6,   2,4,6,  2,4,6)
gammas <- data.frame(setting = factor(rep(1:9, each = 500)), 
                    x = rep(seq(0, 3, length = 500), 9),
                    alpha = rep(alpha, each = 500), beta = rep(beta, each = 500))
gammas <- gammas %>% 
    mutate(y = dgamma(x, shape = alpha, rate = beta))
levels(gammas$setting) <- paste0("Gamma(",alpha,",",beta,")")

  
g <- ggplot(gammas, aes(x = x, y = y)) + 
  lims(x = c(0,1), y = c(0,2.5)) + 
  geom_line() + 
  facet_wrap(~ setting) + 
  labs(x = expression(lambda), y = expression(paste("f(",lambda,")"))) +   scale_x_continuous(breaks = c(0,1,2,3), labels = c("0","1","2","3"))

g
``` 

__Tuning Gamma example__

For our example on spam phone calls, set a prior for $\lambda$ such that $E(\lambda) = 3$ and $\lambda$ most likely is between 2 and 4. You can use `plot_gamma()` function to try out different gamma distributions.

\vspace{5cm}

\newpage
\subsection{Poisson Likelihood}

__The Poisson Model__
Let random variable $X$ be the _number of events_ that occur in a fixed amount of time, where $\lambda$ is the rate at which these events occur.  Then the _dependence_ of $X$ on $\lambda$ can be modeled by the Poisson model with __parameter__ $\lambda$.  In mathematical notation:
 
$X | \lambda \sim \text{Pois}(\lambda)$

Correspondingly, the Poisson model is specified by a conditional pmf:    

$f(x|\lambda) =  \frac{e^{-\lambda}\lambda^x}{x!}\;\; \text{ for } x \in \{0,1,2,\ldots,n\}$

where $f(x|\lambda)$ sums to one across $x$:

$\sum_{x=0}^\infty f(x|\lambda) = 1$

__Likelihood__

\vspace{15cm}

\newpage
\subsection{Gamma-Poisson Conjugacy}

If $f(\lambda) \sim  \text{Gamma}(\alpha, \beta)$   

and if $x_i \sim iid \text{ Poissson}(\lambda) \text{for } i \in 1,\ldots ,n$ 

then $f(\lambda|\vec{x}) \sim \text{Gamma}(\alpha+ \sum x_i, \beta+n)$.

