---
output: 
  pdf_document:
    template: template.tex
---
\section{(Simple)Logistic Regression}

\subsection{Question of the Day}

Y - Accepted to Medical School or Not  
X - GPA  

\subsection{The model}

$logit(\pi) = log(\frac{\pi}{(1-\pi)})=\beta_o+\beta_1X_i$

$Y~\sim\text{Binom}(n,\pi)$

\subsection{Logistic Regression - Frequentist Refresher}

```{r}
library(Stat2Data)
data("MedGPA")
```

```{r}
freq_model <- glm(Acceptance ~ GPA, 
                  data = MedGPA, 
                  family = "binomial")

summary(freq_model)
```

\newpage

\subsection{Logistic Regression Bayesian Way}

Below is the code we had written for a simple linear regression model in `rstan`. How would you modify this code for logistic regression? You do not necessarily need to know exact functions or notation but you can write pseudocode. 

```{r}
normal_regression_model <- "
  data{
  int <lower =0> n;
  vector[n] Y;
  vector[n] X;
  }
  
  parameters{
   real beta_0;
   real beta_1;
   real <lower=0> sigma;
  }
  
  model{
  Y ~ normal(beta_0 +beta_1*X, sigma);
  beta_0 ~ normal(0, 2500);
  beta_1 ~ normal(100, 250);
  sigma ~ exponential(0.00065);
  
  }

"
```

