---
title: "Data visualization review"
author: "Dr. Dogucu"
date: "2020-01-22"
output: 
  xaringan::moon_reader:
    css: slide-styles.css
      
---

layout: true
  
<div class="my-header"></div>

<div class="my-footer"> 
 Copyright &copy; <a href="https://mdogucu.ics.uci.edu">Dr. Mine Dogucu</a>. All Rights Reserved.</div> 

---

```{r message = FALSE, echo = FALSE, warning = FALSE}
library(titanic)
library(tidyverse)

titanic_train <- 
  titanic_train %>% 
  mutate(Survived = case_when(Survived == 0 ~ "Not Survived",
                              Survived == 1 ~ "Survived")) %>% 
  mutate(Pclass = as.factor(Pclass))
```

## Review

Which R functions do you recall from last discussion?

What is the shortcut for `%>%`?

---

## Data 

We will be using `titanic_train` dataset from the `titanic` dataset. 

Glimpse at the data. Does everything look like it should?

---


## Who uses ggplot? Some examples

[BBC](https://bbc.github.io/rcookbook/)  

[FiveThirtyEight](https://fivethirtyeight.com/features/the-rise-of-religiously-inspired-terrorism-in-france/)  

[Master List](http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html#Histogram)  

---

## Three steps of basic visualization with ggplot2 package

1. Make some space (this soon will make sense) using the `ggplot()` function.

2. Draw your axes using `aes()` function.

3. Add your plot, whether it is  histogram, bar plot or something else. We will call these geom objects.

---

## Planning for ggplot

For every plot, 

1) Decide, what variables to use.  
2) Decide which variable is on which axes.  
3) Decide what kind of plot it is.  

---

```{r fig.align = "center", fig.height=6, echo = FALSE}
titanic_train %>% 
  ggplot(aes(x = Fare)) + 
  geom_histogram() +
  theme(text = element_text(size=30))
```

Why are we getting this warning? 


---

```{r fig.align = "center", fig.height=6, echo = FALSE}
titanic_train %>% 
  ggplot(aes(x = Fare)) + 
  geom_histogram(binwidth = 15) +
  theme(text = element_text(size=30))
```
  
What step of the code is binwidth related to?
  
---

## Cheatsheet

Look at the ggplot cheatsheet. Moving on you will attempt at making the plot by finding the geom object from your cheatsheet.
---
<br>
<br>
<br>
```{r fig.align = "center", fig.height=6, echo = FALSE}
titanic_train %>% 
  ggplot(aes(x = Pclass, y = Fare)) +
  geom_boxplot() +
  theme(text = element_text(size=30))
```

---
## You Attempt This First

Make a plot that shows the relationship between age of the passenger and the fare that they paid. Comment whether older passengers bought more expensive tickets.

---

```{r fig.height=6, fig.align='center', echo = FALSE, warning = FALSE}
titanic_train %>% 
  ggplot(aes(x = Age, y = Fare )) +
  geom_point() +
  theme(text = element_text(size=30))
```


---
## Color 

```{r echo = TRUE, eval = FALSE}
titanic_train %>% 
  ggplot(aes(x = Fare )) + 
  geom_histogram(color = "salmon") 
```

```{r fig.height = 5, echo = FALSE}
titanic_train %>% 
  ggplot(aes(x = Fare )) + 
  geom_histogram(color = "salmon") +
  theme(text = element_text(size=30))
```



---

## Fill 

```{r echo = TRUE, eval = FALSE}
titanic_train %>% 
  ggplot(aes(x = Fare)) + 
  geom_histogram(fill = "salmon") 
```

```{r fig.height=5, echo = FALSE}
titanic_train %>% 
  ggplot(aes(x = Fare)) + 
  geom_histogram(fill = "salmon") +
  theme(text = element_text(size=30))
```

---

## Shape

```{r fig.height = 4, fig.align='center', warning = FALSE, eval = FALSE, echo = TRUE}
titanic_train %>% 
  ggplot(aes(x = Age, y = Fare )) +
  geom_point(shape = 4) 
```

```{r fig.height = 4, fig.align='center', warning = FALSE, echo = FALSE}
titanic_train %>% 
  ggplot(aes(x = Age, y = Fare )) +
  geom_point(shape = 4) +
  theme(text = element_text(size=30))
```
---

class: center, inverse
## More colors

<br>
<br>
<br>

[bit.ly/colors-r](bit.ly/colors-r)

---


class: center, inverse
## More shapes

<br>
<br>
<br>

[bit.ly/shapes-r](bit.ly/shapes-r)

---

## Stacked bar plot

Can you guess where the fill argument will go for this stacked bar plot?

```{r fig.height=5, fig.align='center', echo = FALSE}
titanic_train %>% 
  ggplot(aes(x = Pclass, fill = Survived)) +
  geom_bar() +
  theme(text = element_text(size=30))
```





