---
title: "Balance and Sequentiality in Bayesian Analysis"
author: "Dr. Dogucu"
date: "2020-01-22"
output: 
  xaringan::moon_reader:
    css: slide-styles.css
      
---

layout: true
  
<div class="my-header"></div>

<div class="my-footer"> 
 Copyright &copy; <a href="https://mdogucu.ics.uci.edu">Dr. Mine Dogucu</a>. All Rights Reserved.</div> 

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r message = FALSE}
library(titanic)
library(tidyverse)
library(fivethirtyeight)

titanic_train <- 
  titanic_train %>% 
  mutate(Survived = case_when(Survived == 0 ~ "Not Survived",
                              Survived == 1 ~ "Survived")) %>% 
  mutate(Pclass = as.factor(Pclass))
```


---

## Review

Which R functions do you recall from last discussion?

What is the shortcut for `%>%`?

---

## Data 

We will be using ???

---




---



---

## Who uses ggplot? Some examples

[BBC](https://bbc.github.io/rcookbook/)  

[FiveThirtyEight](https://fivethirtyeight.com/features/the-rise-of-religiously-inspired-terrorism-in-france/)  
[Master List](http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html#Histogram)  

---

## Three steps of basic visualization with ggplot2 package

1. Make some space (this soon will make sense) using the `ggplot()` function.

2. Draw your axes using `aes()` function.

3. Add your plot, whether it is  histogram, bar plot or something else. We will call these geom objects.

---

## Planning for ggplot

For every plot, 

1) Decide, what variables to use.  
2) Decide which variable is on which axes.  
3) Decide what kind of plot it is.  




---

class: center, inverse
## More colors

<br>
<br>
<br>

[bit.ly/colors-r](bit.ly/colors-r)

---


class: center, inverse
## More shapes

<br>
<br>
<br>

[bit.ly/shapes-r](bit.ly/shapes-r)

---




